Memory for television episodes preserves event content while introducing new across-event similarities

### 2020-01-06 paper/main.tex
fixed memlab.bib path
updated wordle figure text

Because our analysis framework projects the dynamic video content and participants' recalls onto a shared topic space, and because the dimensions of that space are known (i.e., each topic dimension is a set of weights over words in the vocabulary; Fig.~\topics), we can examine the topic trajectories to understand which specific content was remembered well (or poorly).  For each video event, we can ask: what was the average correlation (across participants) between the video event's topic vector and the closest matching recall event topic vectors from each participant? This yields a single correlation coefficient for each video event, describing how closely participants' recalls of the event tended to reliably capture its content (Fig.~\ref{fig:topics}A).  (We also examined how different comparisons between each video event's topic vector and the corresponding recall event topic vectors related to hand-annotated characterizations of memory performance; see \textit{Supporting Information}).  Given this summary of which events were recalled reliably (or not), we next asked whether the better-remembered or worse-remembered events tended to reflect particular topics.  We computed a weighted average of the topic vectors for each video event, where the weights reflected how reliably each event was recalled.  To visualize the result, we created a ``wordle'' image~\citep{MuelEtal18} where words weighted more heavily by better-remembered topics appear in a larger font (Fig.~\ref{fig:topics}B, green box).  Across the full video, content that reflected topics necessary to convey the central focus of the video (e.g., the names of the two main characters, ``Sherlock" and ``John", and the address of a major recurring location, ``221b Baker Street") were best remembered.  An analogous analysis revealed which themes were poorly remembered.  Here in computing the weighted average over events' topic vectors, we weighted each event in \textit{inverse} proportion to how well it was remembered (Fig.~\ref{fig:topics}B, red box).  The least well-remembered video content reflected information not necessary to conveying the video's ``gist," such as the names of relatively minor characters (e.g., ``Mike," ``Jeffrey," ``Molly," and ``Jimmy") and locations (e.g., ``St. Bartholomew's Hospital").

@@

A similar result emerged from assessing the topic vectors for individual video and recall events (Fig.~\ref{fig:topics}C).  Here, for each of the three best- and worst-remembered video events, we have constructed two wordles: one from the original video event's topic vector (left) and a second from the average recall topic vector for that event (right).  The three best-remembered events (circled in green) correspond to scenes important to the central plot-line (Sherlock and John meeting, Sherlock and John chasing the killer, and the killer calling spying on John in a phone booth).  Meanwhile, the three worst-remembered events (circled in red) reflect various side-stories (John talking to his therapist about the war, a soon-to-be-victim's affair with his assistant, and another soon-to-be-victim leaving a party with his friend) that are not essentially to summarizing the video's narrative.

%Events that reflected topics weighting heavily on characters like ``Sherlock'' and ``John'' (the main characters) and locations like ``221b Baker Street'' (a major recurring location and the address of the flat that Sherlock and John share) were best remembered.  An analogous analysis revealed which themes were poorly remembered.  Here in computing the weighted average over events' topic vectors, we weighted each event in \textit{inverse} proportion to how well it was remembered (Fig.~\ref{fig:topics}B, red box).  This revealed that events with relatively minor characters such as ``Mike,'' ``Jeffrey,'' and ``Molly,'' as well as less-integral plot locations (e.g., ``hospital'' and ``office'') were least well-remembered.  This suggests that what is retained in memory are the major plot elements (i.e., the overall shape of what happened), whereas the more minor details are prone to pruning.

% In addition to constructing overall summaries, assessing the video and recall topic vectors from individual events can provide further insights.  Specifically, for any given event we can construct two wordles: one from the original video event's topic vector, and a second from the average topic vectors produced by all participants' recalls of that event.  We can then examine those wordles visually to gain an intuition for which aspects of the video event were recapitulated in participants' recalls.  Several example wordles are displayed in Figure~\ref{fig:topics}C (wordles from the three best-remembered events are circled in green; wordles from the three worst-remembered events are circled in red).  Using wordles to visually compare the topical content of each video event and the (average) corresponding recall event reveals the specific content from the specific events that is reliably retained in the transformation into memory (green events) or not (red events).

@@

\subsubsection*{Visualizing the video and recall topic trajectories}
We used the UMAP algorithm~\citep{McInHeal18} to project the 100-dimensional topic space onto a two-dimensional space for visualization (Figs.~\ref{fig:trajectory}, \ref{fig:topics}).  To ensure that all of the trajectories were projected onto the \textit{same} lower dimensional space, we computed the low-dimensional embedding on a ``stacked'' matrix created by vertically concatenating the events-by-topics topic proportions matrices for the video and all 17 participants' recalls.  We then divided the rows of the result (a total-number-of-events by two matrix) back into separate matrices for the video topic trajectory and the trajectories for each participant's recalls (Fig.~\ref{fig:trajectory}).  This general approach for discovering a shared low-dimensional embedding for a collections of high-dimensional observations follows \cite{HeusEtal18a}.
% ------ NOTE: ------
% need to add info about embedding optimization
% - Paxton

@@

\bibliography{../../CDL-bibliography/memlab}
%\bibliography{CDL-bibliography/memlab}

>>>>

Because our analysis framework projects the dynamic video content and participants' recalls onto a shared topic space, and because the dimensions of that space are known (i.e., each topic dimension is a set of weights over words in the vocabulary; Fig.~\topics), we can examine the topic trajectories to understand which specific content was remembered well (or poorly).  For each video event, we can ask: what was the average correlation (across participants) between the video event's topic vector and the closest matching recall event topic vectors from each participant? This yields a single correlation coefficient for each video event, describing how closely participants' recalls of the event tended to reliably capture its content (Fig.~\ref{fig:topics}A).  Given this summary of which events were recalled reliably (or not), we next asked whether the better-remembered or worse-remembered events tended to reflect particular topics.  We computed a weighted average of the topic vectors for each video event, where the weights reflected how reliably each event was recalled.  To visualize the result, we created a ``wordle'' image~\citep{MuelEtal18} where words weighted more heavily by better-remembered topics appear in a larger font (Fig.~\ref{fig:topics}B, green box).  Across the full video, content that reflected topics necessary to convey the central focus of the video (e.g., the names of the two main characters, ``Sherlock" and ``John", and the address of a major recurring location, ``221B Baker Street") were best remembered.  An analogous analysis revealed which themes were poorly remembered.  Here in computing the weighted average over events' topic vectors, we weighted each event in \textit{inverse} proportion to how well it was remembered (Fig.~\ref{fig:topics}B, red box).  The least well-remembered video content reflected information not necessary to conveying the video's ``gist," such as the proper names of relatively minor characters (e.g., ``Mike," ``Molly," and ``Lestrade") and locations (e.g., ``St. Bartholomew's Hospital"), as well as the brief, animated clip participants viewed at the beginning of each of the two scan session (involving ``singing" ``cartoon" characters).

@@

A similar result emerged from assessing the topic vectors for individual video and recall events (Fig.~\ref{fig:topics}C).  Here, for each of the three best- and worst-remembered video events, we have constructed two wordles: one from the original video event's topic vector (left) and a second from the average recall topic vector for that event (right).  The three best-remembered events (circled in green) correspond to scenes important to the central plot-line: a mysterious figure spying on John in a phone booth; John and Sherlock discussing the murders in their apartment; and Sherlock laying a trap to catch the murderer.  Meanwhile, the three worst-remembered events (circled in red) reflect scenes that are non-essential to summarizing the narrative's structure: the two appearances of singing cartoon characters; Molly watching as Sherlock beats a corpse in the morgue; and Sherlock noticing evidence of Anderson's and Donovan's affair.

@@

\subsubsection*{Visualizing the video and recall topic trajectories}
We used the UMAP algorithm~\citep{McInHeal18} to project the 100-dimensional topic space onto a two-dimensional space for visualization (Figs.~\ref{fig:trajectory}, \ref{fig:topics}).  To ensure that all of the trajectories were projected onto the \textit{same} lower dimensional space, we computed the low-dimensional embedding on a ``stacked'' matrix created by vertically concatenating the events-by-topics topic proportions matrices for the video and all 17 participants' recalls.  We then divided the rows of the result (a total-number-of-events by two matrix) back into separate matrices for the video topic trajectory and the trajectories for each participant's recalls (Fig.~\ref{fig:trajectory}).  This general approach for discovering a shared low-dimensional embedding for a collections of high-dimensional observations follows \cite{HeusEtal18a}.  Note: for further details on how we created this low-dimensional embedding space, see \textit{Supporting Information}.
% ------ NOTE: ------
% need to add info about embedding optimization
% - Paxton

@@

\bibliography{../../CDL-bibliography/memlab}
% \bibliography{CDL-bibliography/memlab}
